Sketching for Density Estimation

### 2018-12-18 komlos.tex

To invoke Banaszczyk's theorem here we set $x_i \in \R^{d_1d_2}$ to be the flattening of $X_i$ into a vector by concatenating its rows. 
Let $\mathcal K$ be the set of vectors in $\R^{d_1d_2}$ whose stacking (the inverse of flattening) results in a matrix whose spectral norm is less then $c\sqrt{d_1 + d_2}$.
More accurately $\mathcal K = \{x  \in \R^{d_1d_2} |\;\; ||\operatorname{stack}(x)||_2 \le c\sqrt{d_1+d_2}\}$. 
Due to $\operatorname{stack}$ being linear it is obvious that $\mathcal K$ is convex. 
Moreover $\Pr[g \in \mathcal K] \ge 1/2$ due to standard results on the operator norm of random Gaussian matrices \el{add citation \cite{}}.
Banaszczyk's theorem says that there exist signs $\sigma_i$ such that $\sum \sigma_i \operatorname{flatten}(X_i) \in C \mathcal K$.
In other words $\| \sum \sigma_i X_i\| = \| \operatorname{stack}( \sum \sigma_i \operatorname{flatten}(x_i x_i^T)) \| \le C \sqrt{d}$. 
\end{proof}
The algorithm for finding the signs above is, unfortunately, not as simple as with kernels. 
In fact, Banaszczyk's theorem was not constructive and a matching algorithms is a very recent breakthrough result by Bansal, Dadush, Garg, and Lovett \cite{DBLP:conf/stoc/BansalDGL18}.
Using the general framework above gives a matrix covariance coreset of cardinality $O(\sqrt{d}/\eps)$.
Alternatively, we obtain a streaming mergeable algorithm for creating coresets of size $\log^{2}(\min(n, \log(1/\delta)) \sqrt{d}/\eps$.

\el{Selecting $o(1/\eps^2)$ is claimed to be hard by \cite{DBLP:conf/focs/DeshpandeR10} and \cite{DBLP:conf/soda/DeshpandeRVW06}. We need to see if these hold in our setting or only hold for sampling.  Also, \cite{DBLP:conf/soda/GhashamiP14} showed that the impossibility result is limited to frequent directions like settings. Their assumptions don't hold here.} \zk{We can get $1/\eps^2$ though which is better than $\sqrt{d}/\eps$ as long as $\eps > 1/\sqrt{d}$. If $\eps < 1/\sqrt{d}$ then $\sqrt{d}/\eps > d$ - shouldn't it be trivial to get a coreset of size $d$?}

@@

>>>>

To invoke Banaszczyk's theorem here we set $x_i \in \R^{d_1d_2}$ to be the flattening of $X_i$ into a vector by concatenating its rows. 
Let $\mathcal K$ be the set of vectors in $\R^{d_1d_2}$ whose stacking (the inverse of flattening) results in a matrix whose spectral norm is less then $c\sqrt{d_1 + d_2}$.
More accurately $\mathcal K = \{x  \in \R^{d_1d_2} |\;\; ||\operatorname{stack}(x)||_2 \le c\sqrt{d_1+d_2}\}$. 
Due to $\operatorname{stack}$ being linear it is obvious that $\mathcal K$ is convex. 
Moreover $\Pr[g \in \mathcal K] \ge 1/2$ due to standard results on the operator norm of random Gaussian matrices \el{add citation \cite{}}.
Banaszczyk's theorem says that there exist signs $\sigma_i$ such that $\sum \sigma_i \operatorname{flatten}(X_i) \in C \mathcal K$.
In other words $\| \sum \sigma_i X_i\| = \| \operatorname{stack}( \sum \sigma_i \operatorname{flatten}(x_i x_i^T)) \| \le C \sqrt{d}$. 
\end{proof}
The algorithm for finding the signs above is, unfortunately, not as simple as with kernels. 
In fact, Banaszczyk's theorem was not constructive and a matching algorithms is a very recent breakthrough result by Bansal, Dadush, Garg, and Lovett \cite{DBLP:conf/stoc/BansalDGL18}.

Using the general framework above gives a matrix covariance coreset of cardinality $O(\sqrt{d}/\eps)$.
Alternatively, we obtain a streaming mergeable algorithm for creating coresets of size $\log^{2}(\min(n, \log(1/\delta)) \sqrt{d}/\eps$.

\el{Selecting $o(1/\eps^2)$ is claimed to be hard by \cite{DBLP:conf/focs/DeshpandeR10} and \cite{DBLP:conf/soda/DeshpandeRVW06}. We need to see if these hold in our setting or only hold for sampling.  Also, \cite{DBLP:conf/soda/GhashamiP14} showed that the impossibility result is limited to frequent directions like settings. Their assumptions don't hold here.}

@@

In \cite{DBLP:journals/corr/abs-1802-01751} (see paragraph before Lemma 2.2) they manage to obtain a coreset (with equal weights) of size $\sqrt{d \log(1/\eps)}/\eps$ for any kernel with certain properties. The do so by considering a matrix whose rows correspond to all the vectors in an $\eps$-net over the space, and columns correspond to the $n$ data points. Calling that matrix $A$, $A_{q,x} = K(q,x)$. They bound the discrepancy of the matrix defined as $\min_s \|As\|_\infty$ where $s$ is a sign vector by using a bound 
$$ \min_s \|As\|_\infty \leq \sqrt{\log(N)} \|A\|_{\gamma_2} $$
Here, $N$ is the size of the $\eps$ net and the $\gamma_2$ norm is 
$$ \|A\|_{\gamma_2} = \min_{B,C: A=BC^T} \text{max norm of a row of B or C}$$
They prove that for psd kernel $K$, $\|A\|_{\gamma_2} \leq 1$ (in fact it is equal to 1). Now, back to our case. For any fixed power $k$ we know that $K(x,q)=\ip{x,q}^k$ is a psd kernel, hence its corresponding matrix $A_k$ has $\gamma_2$ norm bounded by $1$. Now, the logistic function can be written in a taylor expansion whose coefficients, in absolute value, sum to a constant $C$. It follows that the $A$ matrix for the logistic function has a $\gamma_2$ norm bounded by $C$. The same arguments as in  \cite{DBLP:journals/corr/abs-1802-01751} can now prove that its discrepancy is bounded by $\sqrt{\log(N)} \approx \sqrt{d\log(1/\eps)}$
